---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# BRCore

<!-- badges: start -->

[![R-CMD-check](https://github.com/germs-lab/BRCore/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/germs-lab/BRCore/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

The goal of BRCore is to aid the analysis of the [Inter BRC Microbiome project](https://github.com/germs-lab/interbrc-core-analysis) by providing a set of tools to process and analyze microbial data from Bioenergy Research Centers. ***This is still an experimental package***.

## Installation

Install the development version of `BRCore` from GitHub with:

```{r}
# install.packages("pak")
pak::pak("germs-lab/BRCore")
```

## Step 1: Load Required Libraries and Data

```{r}
library(phyloseq)
library(BRCore)

# Load the esophagus dataset
data(esophagus, package = "phyloseq")
```

## Step 2: Create Phyloseq Object

```{r}
# Get the taxa names from the esophagus dataset
taxa_names <- taxa_names(esophagus)

# Define realistic taxonomy levels
  kingdoms <- c("Bacteria", "Archaea")
  phyla <- c(
    "Firmicutes",
    "Bacteroidetes",
    "Proteobacteria",
    "Actinobacteria",
    "Euryarchaeota"
  )
  classes <- c(
    "Clostridia",
    "Bacteroidia",
    "Gammaproteobacteria",
    "Actinobacteria",
    "Methanobacteria"
  )
  orders <- c(
    "Clostridiales",
    "Bacteroidales",
    "Enterobacterales",
    "Bifidobacteriales",
    "Methanobacteriales"
  )
  families <- c(
    "Lachnospiraceae",
    "Bacteroidaceae",
    "Enterobacteriaceae",
    "Bifidobacteriaceae",
    "Methanobacteriaceae"
  )
  genera <- c("Blautia",
              "Bacteroides",
              "Escherichia",
              "Bifidobacterium",
              "Methanobrevibacter")
  species <- c(
    "Blautia producta",
    "Bacteroides fragilis",
    "Escherichia coli",
    "Bifidobacterium longum",
    "Methanobrevibacter smithii"
  )
  
  # Create a mock taxonomy table
  set.seed(8998); mock_taxonomy_table <- matrix(
    c(
      sample(kingdoms, length(taxa_names), replace = TRUE),
      sample(phyla, length(taxa_names), replace = TRUE),
      sample(classes, length(taxa_names), replace = TRUE),
      sample(orders, length(taxa_names), replace = TRUE),
      sample(families, length(taxa_names), replace = TRUE),
      sample(genera, length(taxa_names), replace = TRUE),
      sample(species, length(taxa_names), replace = TRUE)
    ),
    nrow = length(taxa_names),
    ncol = 7,
    dimnames = list(
      taxa_names,
      c(
        "Kingdom",
        "Phylum",
        "Class",
        "Order",
        "Family",
        "Genus",
        "Species"
      )
    )
  )

# Convert to a taxonomy table object
tax_table <- tax_table(mock_taxonomy_table)

# Add sample metadata to the esophagus dataset
sample_data <- data.frame(
  Sample = sample_names(esophagus),
  Group = sample(c("A", "B"), nsamples(esophagus), replace = TRUE), # Random groups
  row.names = sample_names(esophagus)
)

# Add the taxonomy table and sample metadata to the esophagus dataset
esophagus_with_tax <- merge_phyloseq(esophagus, tax_table, sample_data(sample_data))
```

## Step 2: Run extract_core()

```{r}
# Extract core microbial taxa
core_result <- extract_core(
  physeq = esophagus_with_tax,
  Var = "Group",  # Variable in sample metadata to group by
  method = "increase",  # Method for core taxa selection
  increase_value = 2  # Threshold for core taxa inclusion (2% contribution to Bray-Curtis dissimilarity)
)

# View the results
print(core_result)
```

## Output

The `extract_core()` function returns a list containing the following elements:

-   **core_otus**: A vector of core OTUs identified based on the specified threshold.

-   **bray_curtis_ranked**: A data frame of Bray-Curtis dissimilarity rankings.

-   **otu_rankings**: A data frame of OTU rankings based on occupancy and abundance.

-   **occupancy_abundance**: A data frame of occupancy and abundance values for each OTU.

-   **otu_table**: The OTU table used for analysis.

-   **sample_metadata**: The sample metadata used for analysis.

-   **taxonomy_table**: The taxonomy table used for analysis.
