---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# BRCore

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental) [![](https://www.r-pkg.org/badges/version-last-release/BRCore)](https://www.r-pkg.org/badges/version-last-release/BRCore) [![](https://cranlogs.r-pkg.org/badges/grand-total/BRCore)](https://cran.r-project.org/package=BRCore) [![Codecov test coverage](https://codecov.io/gh/germs-lab/BRCore/graph/badge.svg)](https://app.codecov.io/gh/germs-lab/BRCore) [![R-CMD-check](https://github.com/germs-lab/BRCore/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/germs-lab/BRCore/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

## Overview

BRCore provides a unified framework for identification and ecological interpretation of core microbiomes across time and space, enhancing robustness and reproducibility in microbiome data analysis.

BRCore provides tools for:

**Rarefaction analysis**: Calculate pre-rarefaction metrics and perform rarefactions

-   `add_rarefaction_metrics()`
-   `multi_rarefy()`
-   `plot_variance_propagation()`

**Core microbiome identification**: Identify core microbial taxa using abundance-occupancy distributions

-   `identify_core()`

**Neutral model fitting**: Fit and visualize neutral community models

-   `fit_neutral_model()`
-   `plot_neutral_model()`
-   `scnm.fit()`

**Visualization**: Plot rarefaction diagnostics, abundance-occupancy curves, and core distributions

-   `plot_rarefaction_metrics()`
-   `plot_abundance_occupancy()`
-   `plot_core_distribution()`
-   `plot_identified_core()`

## Installation

Install the latest *stable* version of BRCore from CRAN with:

```{r}
install.packages("BRCore")
```

Install the *development* version of BRCore from GitHub with:

Install the *development* version of BRCore from GitHub with:
```{r}
# install.packages("pak")
pak::pak("germs-lab/BRCore")
```

***Note:*** *`pak` handles dependencies automatically.*

## Quick Start

```{r}
library(BRCore)
library(phyloseq)

# Load example data
data("bcse", package = "BRCore")

# Add rarefaction metrics
bcse_metrics <- add_rarefaction_metrics(data = bcse)

# Perform rarefaction
bcse_rarefied_list <- multi_rarefy(
  physeq_obj = bcse,
  depth_level = 1000,
  num_iter = 3,
  set_seed = 7642
)

# Update phyloseq object with rarefied data
bcse_rare_single <- update_otu_table(
  physeq_obj = bcse, 
  rarefied_otus = bcse_rarefied_list, 
  iteration = 2) # Your preffered iteration can be used here

# Identify core microbiome

# With a single iteration of rarefaction
bcse_core <- identify_core(
  physeq_obj = bcse_rare_single,
  priority_var = "Crop",
  increase_value = 0.02,
  seed = 2134
)


# With multiple iterations of rarefaction
bcse_core_multi <- identify_core(
  physeq_obj = bcse, 
  rarefied_list = bcse_rarefied_list
  priority_var = "Crop",
  increase_value = 0.02,
  depth_level = 1000,
  num_iter = 10,
  seed = 2134
)

# Visualize abundance-occupancy distribution
plot_abundance_occupancy(core_result = bcse_core, core_set = "increase")

# Fit and plot neutral model
bcse_neutral <- fit_neutral_model(
  otu_table = bcse_core$otu_table,
  core_set = bcse_core$increase_core,
  abundance_occupancy = bcse_core$abundance_occupancy
)

plot_neutral_model(bcse_neutral)
```

## Documentation

For detailed examples and use cases, see the package vignette:

```{r}
vignette("BRCore-vignette", package = "BRCore")
```

## Contributing

Contributions to BRCore are welcome! Please see the [CONTRIBUTING.md](https://github.com/germs-lab/BRCore/blob/main/.github/CONTRIBUTING.md) file for guidelines on how to contribute.

## Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](https://github.com/germs-lab/BRCore/blob/main/.github/CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.
